Title goes here

### 2016-10-23 paper/paper.tex
PAPER: add content

\begin{abstract}
Abstract goes here.  
\end{abstract}

@@

@@

@@

@@

% TODO insert picture of route coloring here

@@

% also add results of the different phases in terms of file size. for example,
% how much the data shrinks from raw avro to position messages only.

@@

@@

>>>>

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.28]{imgs/4b7fad-altitudes.png}
  \caption{Altitude values of the aircraft with ICAO 4b7fad, during September
    20th, 2016. It can be seen that the data points disappear when the aircraft
    approaches the ground. Also, in the group of points of the left, the
    information on the first part of the flight is entirely missing, and the
    first point available has timestamp 06:46 am. This is almost
    certainly due to the airplane traveling through a poorly covered area.}
  \label{fig:alts}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.28]{imgs/4b7fad-altitudes-grouped.png}
  \caption{Altitude values of the aircraft with ICAO 4b7fad, during September
    20th, 2016, as grouped by the flight detection algorithm.}
  \label{fig:alts}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.4]{imgs/amsterdam-vatry-co2.png}
  \caption{Visualization of CO2 consumption for the standard route from
    Amsterdam to Vatry. It can be seen that the route is mostly green for the
    first part. It then shifts to red in the last part, where it takes a detour
    from the straight-line route.}
  \label{fig:amsterdam-vatry}
\end{figure}

@@

@@

When working of large-scale datasets, it is essential to understand the
available data and try to extract only relevant information from it. The
subdivision of our algorithm in different phases is not only a convenient
logical partitioning of operations, but it is essential to transform the dataset
into smaller datasets, tailored to our needs. It is useful to do this
transformation gradually, step by step, so that single intermediate phases can
be reiterated without affecting the previous ones.

The whole algorithm has been implemented as a Scala + Spark project, that has
been deployed to the Hadoop cluster. We developed every phase of the algorithm
as a separate executable Scala class, allowing us to run each phase separately.

The position extraction phase was run on the entire dataset of raw AVRO-encoded
messages, and it was, and one could expect, the most time consuming. The entire
execution took about 48 hours. As a result of this phase, we were able to shrink
the data from about 85 GB per day to about 10 GB of position points per day.
This of course made the next phases of the algorithm more tractable. Position
data has then been processed by the flight detection algorithm, in about 24
hours. This phase, as we already explained, throws away a lot of data. Also,
flights are represented and saved to disk as sequences of sectors, where each
flight it composed of 20--50 sectors. This gives a very succint representation,
and we were able to store the output of this phase in just about 4 MB per day.
The small size of the flights data allowed us to run the route clustering
algorithm and all the additional JSON files generation for the visualization in
under 30 minutes. With such small execution time, we were able to test the
results of the algorithm several times, and improve its performance and results.

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.4]{imgs/gatwick-schiphol.png}
  \caption{Results of standard (blue) and diverging (red) routes detection for
    flights from London Gatwick to Amsterdam Schiphol.}
  \label{fig:gatwick-schiphol}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.35]{imgs/schiphol-linate.png}
  \caption{Results of standard (blue) and diverging (red) routes detection for
    flights from Amsterdam Schiphol to Milano Linate. It can be seen that the
    diverging route on the right is very similar to the standard route, but the
    clustering algorithm fails to classify it as such.}
  \label{fig:schiphol-linate}
\end{figure}
